---
title: "bcrange2020.3"
author: "jeremy horowitz"
date: "February 27, 2020"
output: html_document
---

```{r}
library(tidyverse)
library(plotly)
library(cowplot)
library(sp)
library(quantreg)
```


plotting all lat on one plot for atl
```{r}

attach(atl_ranges3)
atl_subset<-subset(atl_ranges3,ocean=="atl")
atl_lat_all<-atl_subset[ atl_subset$type %in% c("lat_range","lat_planar_range","lat_geog_range"),]

#remove bounding points
atl_lat_all<-subset(atl_lat_all,!(species=="bound"))

#log the dependent variable:
#atl_lat_all<-atl_lat_all %>% mutate(area=log10(area))

write.csv(atl_lat_all,"output/atl_lat_all.csv")

attach(atl_lat_all)

atl_lat_all$type<-factor(atl_lat_all$type,levels=c("lat_range","lat_planar_range","lat_geog_range"))

lat_plot_atl<-ggplot(data=atl_lat_all,
       aes(mid,area,color=type))+geom_point()+
  geom_point(aes(y=area))+
  geom_smooth(se=F,method=loess,size=2)+
  theme_classic()+
  xlim(-80,80)+
  scale_y_continuous(trans='log10')+
  xlab("Mid-latitude (decimal degree)")+
  ylab("range extent m/m2")+
  geom_point(aes(x=-80, y=0), colour="black")+
  geom_point(aes(x=-40, y=1.445025e+14), colour="black")+
  geom_point(aes(x=-0, y=4.295588e+14), colour="black")+
  geom_point(aes(x=40, y=2.323491e+14), colour="black")+
  geom_point(aes(x=80, y=0), colour="black")+
  geom_point(aes(x=-40, y=2.890000e+14), colour="black")+
  geom_point(aes(x=-0, y=8.590000e+14), colour="black")+
  geom_point(aes(x=40, y=4.650000e+14), colour="black")+
  geom_point(aes(x=-40, y=1.518389e+07), colour="black")+
  geom_point(aes(x=0, y=3.036779e+07), colour="black")+
  geom_point(aes(x=40, y=1.518389e+07), colour="black")
  
lat_plot_atl

ggsave(filename="lat_plot_atl.tiff",path="figs/")



```
no species occur with mid-latitudes <-30 dec deg lat. the largest ranged species occur between -30 and 0 degrees lat, and then ranges decrease in area northerly. there is a dip in species range size at 30 deg lat. this can be explaained by the relatively large number of small-ranged, possibly endemic species occuring in the caribbean. there are also relatively large ranged species at this latitude but the caribbean species are greter in richness and therefore modify the trend. the trend does not change depending on the way ranges are considered, which means that latitude extent alone is sufficent to represent range area vs mid-lat relationships.

caribb=seperate biome with separate evolutionary history

```{r}
attach(atl_ranges3)
pac_subset<-subset(atl_ranges3,ocean=="pac")
pac_lat_all<-pac_subset[pac_subset$type %in% c("lat_range","lat_planar_range","lat_geog_range"),]

#remove bounding points
pac_lat_all<-subset(pac_lat_all,!(species=="bound"))

#log the dependent variable:
#pac_lat_all<-atl_lat_all %>% mutate(area=log10(area))

write.csv(pac_lat_all,"output/pac_lat_all.csv")

attach(pac_lat_all)

pac_lat_all$type<-factor(pac_lat_all$type,levels=c("lat_range","lat_planar_range","lat_geog_range"))

lat_plot_pac<-ggplot(data=pac_lat_all,
       aes(mid,area,color=type))+geom_point()+
  geom_point(aes(y=area))+
  geom_smooth(se=F,method=loess,size=2)+
  theme_classic()+
  xlim(-80,80)+
  scale_y_continuous(trans='log10')+
  xlab("Mid-latitude (decimal degree)")+
  ylab("range extent m/m2")+
  geom_point(aes(x=-80, y=0), colour="black")+
  geom_point(aes(x=-45.000000, y=2.368199e+14), colour="black")+
  geom_point(aes(x=-10, y=4.185974e+14), colour="black")+
  geom_point(aes(x=30, y=1.603900e+14), colour="black")+
  geom_point(aes(x=70, y=0), colour="black")+
  geom_point(aes(x=-45, y=4.737704e+14), colour="black")+
  geom_point(aes(x=-10, y=8.372510e+14), colour="black")+
  geom_point(aes(x=30, y=3.208763e+14), colour="black")+
  geom_point(aes(x=70, y=0), colour="black")+
  geom_point(aes(x=-45, y=1.401611e+07), colour="black")+
  geom_point(aes(x=-00, y=2.605160e+07), colour="black")+
    geom_point(aes(x=30, y=1.192319e+07), colour="black")

  
lat_plot_pac

ggsave(filename="lat_plot_pac.tiff",path="figs/")
```
relationship shape roughly relates to mid-domain effect, slightly right skewed with largest ranges being at ~-20deg. a small dip can also be seen at equator. cite paper that shows bimodal relationshpi between richness or range size in lat. shape is = for each range type meaning lat extent is enough to depict trends. relate this to map.


atl long all
```{r}

attach(atl_ranges3)
atl_subset<-subset(atl_ranges3,ocean=="atl")
atl_long_all<-atl_subset[ atl_subset$type %in% c("long_range","long_planar_range","long_geog_range"),]

#remove bounding points
atl_long_all<-subset(atl_long_all,!(species=="bound"))

#log the dependent variable:
#atl_long_all<-atl_long_all %>% mutate(area=log10(area))

write.csv(atl_long_all,"output/atl_long_all.csv")

attach(atl_long_all)

atl_long_all$type<-factor(atl_long_all$type,levels=c("long_range","long_planar_range","long_geog_range"))

long_plot_atl<-ggplot(data=atl_long_all,
       aes(mid,area,color=type))+geom_point()+
  geom_point(aes(y=area))+
  geom_smooth(se=F,method=loess,size=2)+
  theme_classic()+
  xlim(-110,45)+
  scale_y_continuous(trans='log10')+
  xlab("Mid-latitude (decimal degree)")+
  ylab("range extent m/m2")+
  geom_point(aes(x=-106, y=0), colour="black")+
  geom_point(aes(x=-68.375000, y=2.094712e+14), colour="black")+
  geom_point(aes(x=-30.750000, y=4.295588e+14), colour="black")+
  geom_point(aes(x=6.875000, y=2.159317e+14), colour="black")+
  geom_point(aes(x=44.5, y=0), colour="black")+
  geom_point(aes(x=-68.375000, y=4.190000e+14), colour="black")+
  geom_point(aes(x=-30.750000, y=8.590000e+14), colour="black")+
  geom_point(aes(x=6.875000, y=4.300000e+14), colour="black")+
  geom_point(aes(x=-68.375000, y=8.404622e+06), colour="black")+
  geom_point(aes(x=-30.750000, y=1.680924e+07), colour="black")+
  geom_point(aes(x=6.875000, y=8.404622e+06), colour="black")
  
long_plot_atl

ggsave(filename="long_plot_atl.tiff",path="figs/")

```
relationships relate to mid domain effect. trend shapes are = for each range type. only long is required to depict relationship.

pac long
```{r}

attach(atl_ranges3)
pac_subset<-subset(atl_ranges3,ocean=="pac")
pac_long_all<-pac_subset[pac_subset$type %in% c("long_range","long_planar_range","long_geog_range"),]

#remove bounding points
pac_long_all<-subset(pac_long_all,!(species=="bound"))

#log the dependent variable:
#pac_long_all<-pac_long_all %>% mutate(area=log10(area))

write.csv(pac_long_all,"output/pac_long_all.csv")

attach(pac_long_all)

pac_long_all$type<-factor(pac_long_all$type,levels=c("long_range","long_planar_range","long_geog_range"))

long_plot_pac<-ggplot(data=pac_long_all,
       aes(mid_long_adjust,area,color=type))+geom_point()+
  geom_point(aes(y=area))+
  geom_smooth(se=F,method=loess,size=2)+
  theme_classic()+
  #xlim(-110,45)+
  scale_y_continuous(trans='log10')+
  xlab("Mid-latitude (decimal degree)")+
  ylab("range extent m/m2")+
  geom_point(aes(x=1, y=0), colour="black")+
  geom_point(aes(x=60.83, y=2.568425e+14), colour="black")+
  geom_point(aes(x=91.620, y=4.185974e+14), colour="black")+
  geom_point(aes(x=144.500, y=1.507696e+14), colour="black")+
  geom_point(aes(x=192.000, y=0), colour="black")+
  geom_point(aes(x=60.830, y=5.138604e+14), colour="black")+
  geom_point(aes(x=91.620, y=8.372510e+14), colour="black")+
  geom_point(aes(x=144.500, y=3.016625e+14), colour="black")+
  geom_point(aes(x=192.000, y=0), colour="black")+
  geom_point(aes(x=-1, y=0), colour="black")+
  geom_point(aes(x=60.830, y=1.202251e+07), colour="black")+
  geom_point(aes(x=91.620, y=2.103938e+07), colour="black")+
  geom_point(aes(x=144.500, y=9.016879e+06), colour="black")

  
long_plot_pac

ggsave(filename="long_plot_pac.tiff",path="figs/")
```
get the adj numbers. explain the dip at ~midway. shape = for each type. long is enough.

```{r}
attach(atl_ranges3)
pac_subset<-subset(atl_ranges3,ocean=="pac")
pac_depth_all<-pac_subset[pac_subset$type %in% c("depth_range","depth_planar_range"),]

#remove bounding points
pac_depth_all<-subset(pac_depth_all,!(species=="bound"))

#log the dependent variable:
#pac_long_all<-pac_long_all %>% mutate(area=log10(area))

write.csv(pac_depth_all,"output/pac_depth_all.csv")

attach(pac_depth_all)

pac_depth_all$type<-factor(pac_depth_all$type,levels=c("depth_range","depth_planar_range"))

depth_plot_pac<-ggplot(data=pac_depth_all,
       aes(mid,area,color=type))+geom_point()+
  geom_point(aes(y=area))+
  geom_smooth(se=F,method=loess,size=2)+
  theme_classic()+
  #xlim(-110,45)+
  scale_y_continuous(trans='log10')+
  xlab("Mid-latitude (decimal degree)")+
  ylab("range extent m/m2")+
  geom_point(aes(x=0, y=0), colour="black")+
  geom_point(aes(x=60.83, y=6.279183e+14), colour="black")+
  geom_point(aes(x=5.0, y=7.433056e+14), colour="black")+
  geom_point(aes(x=2260.00, y=8.372495e+14), colour="black")+
  geom_point(aes(x=4520.00, y=8.372510e+14), colour="black")+
  geom_point(aes(x=6780.00, y=5.392336e+14), colour="black")+
  geom_point(aes(x=9040.25, y=0), colour="black")+
  geom_point(aes(x=2260., y=4.520000e+03), colour="black")+
  geom_point(aes(x=4520.00, y=9.040000e+03), colour="black")+
  geom_point(aes(x=6780.00, y=4.520000e+03), colour="black")+
  geom_point(aes(x=9040.00, y=0), colour="black")

  
depth_plot_pac

ggsave(filename="depth_plot_pac.tiff",path="figs/")


```
shapes are different because 2d, including continent edges will result in a large range. 1d shape is ___.range relates perfect to mid domain. shapes are different betwen types but for both, ranges increase with mid lat to mid-lat of ocean, then decreases.

```{r}
attach(atl_ranges3)
atl_subset<-subset(atl_ranges3,ocean=="atl")
atl_depth_all<-atl_subset[atl_subset$type %in% c("depth_range","depth_planar_range"),]

#remove bounding points
atl_depth_all<-subset(atl_depth_all,!(species=="bound"))

#log the dependent variable:
#pac_long_all<-pac_long_all %>% mutate(area=log10(area))

write.csv(atl_depth_all,"output/atl_depth_all.csv")

attach(atl_depth_all)

atl_depth_all$type<-factor(atl_depth_all$type,levels=c("depth_range","depth_planar_range"))

depth_plot_atl<-ggplot(data=atl_depth_all,
       aes(mid,area,color=type))+geom_point()+
  geom_point(aes(y=area))+
  geom_smooth(se=F,method=loess,size=2)+
  theme_classic()+
  #xlim(-110,45)+
  scale_y_continuous(trans='log10')+
  xlab("Mid-latitude (decimal degree)")+
  ylab("range extent m/m2")+
  geom_point(aes(x=0, y=0), colour="black")+
  geom_point(aes(x=2.50, y=6.130000e+13), colour="black")+
  geom_point(aes(x=5.0, y=7.620000e+14), colour="black")+
  geom_point(aes(x=1979.00, y=8.590000e+14), colour="black")+
  geom_point(aes(x=3958.50, y=8.590000e+14), colour="black")+
  geom_point(aes(x=5937.75, y=3.790000e+14), colour="black")+
  geom_point(aes(x=7917.00, y=0), colour="black")+
  geom_point(aes(x=1979.00, y=3.958000e+03), colour="black")+
  geom_point(aes(x=3958.50, y=7.917000e+03), colour="black")+
  geom_point(aes(x=5937.75, y=3.958000e+03), colour="black")
  
depth_plot_atl

ggsave(filename="depth_plot_atl.tiff",path="figs/")

```
shapes are different because 2d, including continent edges will result in a large range. 1d shape is ___.range relates perfect to mid domain. shapes are different betwen types but for both, ranges increase with mid lat to mid-lat of ocean, then decreases.

atl lat richness
```{r}
attach(atl_lat_rich_combine4)


atl_lat_rich_combine4plot<-ggplot(data=atl_lat_rich_combine4,aes(lat,rich,colour=type))+
geom_line()+
#geom_smooth(se=F,method=loess)+
xlab(label="lat (dec deg)")+
ylab("Richness")+
theme_classic()
atl_lat_rich_combine4plot

#to save figure 
ggsave(filename="atl_lat_rich_combine4plot.tiff",path="figs/")

```

pac lat rich
```{r}
attach(pac_temp_combine4)
pac_temp_combine4plot<-ggplot(data=pac_temp_combine4,aes(lat,rich,colour=type))+
geom_line()+
#geom_smooth(se=F,method=loess)+
xlab(label="Lat (dec deg)")+
ylab("Richness")+
theme_classic()
pac_temp_combine4plot

#to save figure 
ggsave(filename="pac_temp_combine4plot.tiff",path="figs/")

```

pac long richness
```{r}
attach(pac_temp_combine2_long_adj)

ggplot(pac_temp_combine2_long_adj,aes(x_adjust))+
geom_line(aes(y=richness_long,color="richness_long"),size=1)+  
geom_line(aes(y=max_rich,color="richness_geog"),size=1)+
geom_line(aes(y=rich_2d,color="rich_2d"),size=1)+
xlab(label="longitude (degrees)")+
ylab("Richness")+
theme_classic()


ggsave(filename="pac_long_rich.tiff",path="figs/")

```



atl long richness
```{r}

attach(atl_long_rich_combine4)
atl_long_rich_combine4plot<-ggplot(data=atl_long_rich_combine4,aes(long,rich,colour=type))+
geom_line()+
#geom_smooth(se=F,method=loess)+
xlab(label="long (dec deg)")+
ylab("Richness")+
theme_classic()
atl_long_rich_combine4plot

#to save figure 
ggsave(filename="atl_long_rich_combine4plot.tiff",path="figs/")

```

pac long rich
```{r}

attach(pac_temp_combine2_long_adj3)
pac_temp_combine2_long_adj3plot<-ggplot(data=pac_temp_combine2_long_adj3,aes(long_adj,rich,colour=type))+
geom_line()+
#geom_smooth(se=F,method=loess)+
xlab(label="Long (dec deg)")+
ylab("Richness")+
theme_classic()
pac_temp_combine2_long_adj3plot

#to save figure 
ggsave(filename="pac_temp_combine2_long_adj3plot.tiff",path="figs/")
```


pac depth rich
```{r}

attach(pac_depth_rich3)
pac_depth_rich3plot<-ggplot(data=pac_depth_rich3,aes(depth,rich,colour=type))+
geom_line()+
#geom_smooth(se=F,method=loess)+
xlab(label="Depth (m)")+
ylab("Richness")+
theme_classic()
pac_depth_rich3plot

#to save figure 
ggsave(filename="pac_depth_rich3plot.tiff",path="figs/")

```


atl depth rich
```{r}

#in excel, changed names of columns for atl_depth_rich3

attach(atl_depth_rich3)

atl_depth_rich3plot<-ggplot(data=atl_depth_rich3,aes(depth,rich,colour=type))+
geom_line()+
#geom_smooth(se=F,method=loess)+
xlab(label="Depth (m)")+
ylab("Richness")+
theme_classic()
atl_depth_rich3plot

#to save figure 
ggsave(filename="atl_depth_rich3plot.tiff",path="figs/")

```
